Formal Modelling of Predator Preferences using Molecular Gut-Content Analysis

### 2015-05-18 paper/main.tex
some concluding words for the discussion

With point estimates of $c_{st}$ under the model $c_{st} = \lambda_{st}/\gamma_{st}$, we can test any number of linear contrasts.  For instance, the hypotheses $c_{1t} = c_{2t}$, for $t \in \{1, \ldots, 18\}$ state that wolf spiders equally prefer the orders Diptera and Collembola at each of the $18$ time points.  Using a level of significance of $0.05$, and after making a Bonferroni multiple comparisons adjustment, the data can not say that the two prey are differently preferred in October, November, and December of $2011$ and for March and July of $2012$.  {\color{red} What does this tell us/someone? Just a quick sentence.}

@@

A more recent method, by \citet{Agusti:2003}, esimates predator's preferences with unobserved count data using a Monte Carlo approach, the bootstrap to be more accurate.  This method deserves credit for its attempt to do away with distributional assumptions, however we find that the lack of a distribution causes some unresolved problems.  \citet{Agusti:2003} suggest resampling the data to estimate proprotions of prey species present in the mesocosms\footnote{Mesocosm was defined by \citet{Agusti:2003} and is there assumed representative of the predator's habitat.} and in predators' guts.  A seemingly intuitive null hypothesis to test random feeding states that pairs of proportions, relative to a given prey species, are equal.  For instance, if the proportion of prey species $j$ in the mesocosm is denoted $\pi_j$, and $p_j$ denotes the proportion of prey species $j$ in the predator's gut, then their null hypothesis is written $\pi_j = p_j, \, \forall j$.  Since the $\pi_j$s depend on the unknown number of prey species consumed, this null hypothesis is not easily tested.  Even if one assumed known the distribution of prey species consumed, this assumed distribution would depend on unkown parameters, thus only complicating any derived null hypothesis further.  Exactly what \citet{Agusti:2003} are testing in their paper is unclear.

The model presented here is a formal statistical model, which enables clear and intuitive hypothesis testing of multiple prey species across an array of time points.  Our methods, via a likelihood ratio test, offer statistically powerful conclusions about the hypotheses of interest.  Robust simulations provided insight into the statistical properties of our model.  Moreover, we can estimate parameters of interest even when the predator's gut contents are not fully observed.

@@

{\color{red} So what does our model offer ecologists? Multiple prey species and time points, contrasts, unobserved count data.  Just a quick paragraph. It should also mention our package.}

>>>>

With point estimates of $c_{st}$ under the model $c_{st} = \lambda_{st}/\gamma_{st}$, we can test any number of linear contrasts.  For instance, the hypotheses $c_{1t} = c_{2t}$, for $t \in \{1, \ldots, 18\}$ state that wolf spiders equally prefer the orders Diptera and Collembola at each of the $18$ time points.  Using a level of significance of $0.05$, and after making a Bonferroni multiple comparisons adjustment, the data can not say that the two prey are differently preferred in October, November, and December of $2011$ and for March and July of $2012$.

@@

A more recent method for testing prey preferences was presenteby by Agusti et al (2003).  This method attempts to test the hypothesis that individuals are eating prey at random using a test based on Monte Carlo simulation (more specificially the bootstrap).  Unfortunately, we believe that the method described is flawed specifically because it does not attempt to model the distribution of the number of prey items eaten, as we have done.  The null hypothesis tested by Agusti et all (2003) is that the proportion of prey species $j=1,\ldots,J$ in the mesocosm, denoted $\pi_j$, is equal to the proportion of prey species $j$ in the predator's guts, denoted by $p_j$, at all of the sampling times.  This is tested by repeatedly simulating the gut contents of spiders in each sample under this hypothesis and then comparing summary statistics computed from the observed data and the simulated data to generate a $p$-value.  The problem with this approach is that the relationship between $\pi_j$ and $p_j$ depends on the number of prey items eaten by each predator.  If each predator had eaten exactly one prey item then the null hypothesis of Agusti et al (2003) would be equivalent to random sampling of prey.  However, the relationship between $\pi_j$ and $p_j$ is more complicated when predators can have more than one prey item in their gut.  Given that a predator has $m$ prey items in its gut, randomly sampled from the mesocosm, the marginal probability that at least one item from prey species $j$ is present is $1-(1-\pi_j)^m$.  The marginal probability that an individual has as least one item from prey species $j$ in its gut is then $\sum_{m=1}^{\infty} (1-(1-\pi_j)^m)P(m)$, where $P(m)$ is the probability that a predator has a total of $m$ species in its gut.  The presence of species in a predators gut contents also cannot be simulated without modeling the number of items in the gut for exactly the same reason.  This shows the importance of the distribution of prey items in the gut in testing the hypothesis of random sampling and because of this it is unclear what hypothesis is being tested by Agusti et al (2003).  Monte Carlo simulation methods similar to this could be implemented using the output from our model, but these would be computationally intensive in comparison to the likelihood ratio tests which can be computed almost immediately.

@@

Our methods offer ecologists a formal statistical framework to model and test predators prey preferences across an array of time points and prey species.  Because we build from intuitive assumptions and well behaved statistical distributions, we were able to extend our methods to unobserved count data.  For both cases, observed or unobserved count data, a computationally efficient likelihood ratio test provides the most powerful test one can hope for.  No methods before have offered such abilities.  These methods are available to researchers via the \texttt{R} package \texttt{spiders}.
